Value-Suppressing Uncertainty Maps

### 2017-03-21 paper/template.tex

@@

	Uncertainty information is a vital part of many datasets, and should be accurately conveyed to the analyst.
	Directly integrating uncertainty information with data requires that designers construct bivariate maps that encode value and uncertainty simultaneously. The design of such bivariate maps can be difficult, and the impact of different choices of encodings on decision-making under uncertainty understudied.
    In this paper, we present a category of bivariate maps for uncertainty, ``value-suppressing uncertainty maps,'' (VSUMS). As opposed to traditional bivariate maps, VSUMS make the informed decision to alias together values when uncertainty is high, in order to afford greater discriminability when uncertainty is low. That is, certain data is intrinsically easier to visually disambiguate than uncertain data. 
    We present several examples of these maps, as well as a crowd-sourced evaluation showing X. We conclude that Y.

	%However, in thematic maps, strong visual variables such as position and hue are already reserved to encode location and value.

	%Remaining visual variables with semiotic connection to uncertainty, such as transparency and saturation, are comparatively more difficult to discriminate, and can interfere with color discrimination.

	%Uncertainty, when it is directly encoded at all, is often visualized using visual variables with lower discriminability, or relegated to tooltips or separate charts.

@@

Uncertainty is an inescapable component of collecting, presenting, and using data. A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty. A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data. That is, it should be difficult to discount or ignore the uncertainty in a dataset. A designer of a visualization could instantiate this principle through the simultaneous presentation of uncertainty and value. This simultaneous presentation necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.

In this paper, we present \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps. VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases. Rather than a square bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value. Figure X shows an example of a VSUM, compared to a more traditional bivariate map. In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate VSUMs in a crowd-sourced experiment.

@@

% What this section needs to do:
% State of the art in uncertainty vis, with heavy emphasis on MacEachren & co. 
% Bivariate maps are hard!
% Binning colormaps is de rigueur
% The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability

\subsection{Visual Variables for Representing Uncertainty}

MacEachren et al. \cite{maceachren2012visual} evaluate a number of visual variables for representing uncertainty. 
\subsection{Interactions Between Visual Variables}

>>>>

%MC Packages and commands
\usepackage{subcaption}
\newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}\\}

@@

	Uncertainty is a vital part of many datasets, and should be conveyed accurately to the analyst.
	Directly incorporating uncertainty requires that designers construct bivariate mappings that encode value and uncertainty simultaneously.
	In this paper, we present a technique for generating such bivariate maps for uncertainty, ``Value-suppressing uncertainty maps'' (VSUMs).
	As opposed to traditional bivariate maps, VSUMs make the informed choice to alias together values when uncertainty is high, in order to afford greater discriminability between categories when uncertainty is low. That is, certain data is intrinsically easier to visually disambiguate than uncertain data.
	Values with low uncertainty can be more precisely compared, whereas similar values with higher uncertainty will be harder to differentiate in accordance with their variability.
	By making informed decisions about where to create category bins, VSUMs make better use of the limited budget of discriminable marks in bivariate maps, and dissuade analysts from making decisions based on data with sufficiently high uncertainty.
	We present several examples of VSUMs, as well as a crowd-sourced evaluation showing that, compared to more traditional bivariate maps, VSUMs promote decision-making that places greater weight on uncertainty information compared to traditional bivariate maps.

@@

Uncertainty is an inescapable component of collecting, presenting, and using data. A common goal in the communication of uncertainty is promoting \textbf{uncertainty-aware decisions}--- that is, the audience should be aware of the risks and rewards of certain decisions, modulate their confidence in their conclusions, and perhaps refrain from making a decision at all if there is too much uncertainty. In general, a goal of designers of visualizations of uncertain data is the \emph{information fusion} of data and uncertainty~\cite{riveiro2007evaluation}. A way that designers can contribute to this goal is by ensuring that uncertainty information is \textbf{well-integrated} with the rest of the data. That is, it should be difficult to discount or ignore the uncertainty in a dataset. Simultaneous presentation of uncertainty and value necessitates the construction of a bivariate map--- a relation, in terms of visual variables, between a 2-tuple $(value$,$uncertainty)$ and a set of marks.

In this paper, we present \textbf{Value-Suppressing Uncertainty Maps} (VSUMs) for integrating data and uncertainty information in thematic maps. VSUMs intentionally alias together data values with high uncertainty, affording greater discriminability as certainty increases. Rather than a square bivariate map, with unique outputs for each relevant combination of data and uncertainty, VSUMs can be thought of as arcs: as uncertainty increases, values are mapped to smaller and smaller sets of outputs, culminating in a singularity where all inputs are mapped to an identical, highly uncertain mark regardless of their data value. Figure X shows an example of a VSUM, compared to a more traditional bivariate map. In this work, we describe the motivations behind the use of VSUMs, examples of their utility for decision-making under uncertainty, and validate that VSUMs in a crowd-sourced experiment.

@@

% What this section needs to do:
% State of the art in uncertainty vis, with heavy emphasis on MacEachren & co. 
% Bivariate maps are hard!
% Binning colormaps is de rigueur
% The variables we'd want to use for uncertainty (like value or alpha or size) mess with color discriminability
Despite the acknowledged importance of uncertainty in understanding data, explicit representation of uncertainty often missing from visualizations~\cite{boukhelifa2009uncertainty}. The absence of

The presentation Broadlie et al.~\cite{brodlie2012review}
\subsection{Visual Variables for Representing Uncertainty}

MacEachren et al.~\cite{maceachren2012visual} evaluate a number of visual variables for representing uncertainty. 
\subsection{Interactions Between Visual Variables}
